Lesson 6
========================================================

### Welcome
Notes:

***

### Scatterplot Review

```{r Scatterplot Review}

library(tidyverse)
ggplot(data=subset(diamonds,carat<quantile(carat,.99)&price<quantile(price,.99)),aes(x=carat,y=price))+
  geom_point(fill=I('#F79420'),color=I('black'),shape=21)+
  stat_smooth(method='lm')
```

***

### Price and Carat Relationship
Response: price and carat are likely positively correlated, but it is not a linear relationship necessarily. 

***

### Frances Gerety
Notes:

#### A diamonds is forever


***

### The Rise of Diamonds
Notes: most successful ad campaign of the 20th century

***

### ggpairs Function
Notes:

```{r ggpairs Function}
# install these if necessary
# install.packages('GGally')
# install.packages('scales')
# install.packages('memisc')
# install.packages('lattice')
# install.packages('MASS')
# install.packages('car')
# install.packages('reshape')
# install.packages('plyr')

# load the ggplot graphics package and the others
library(ggplot2)
library(GGally)
library(scales)
library(memisc)

# sample 10,000 diamonds from the data set
set.seed(20022012)
diamond_samp <- diamonds[sample(1:length(diamonds$price), 10000), ]
ggpairs(diamond_samp,
  lower = list(continuous = wrap("points", shape = I('.'))),
  upper = list(combo = wrap("box", outlier.shape = I('.'))))
```

What are some things you notice in the ggpairs output?
Response: relationships between price and color and price and clarity. also interesting long tails on price and carat, and bimodal distributions of dimensions.

weight is proportional to volume, by density.

***

### The Demand of Diamonds
Notes:higher demand for less expensive diamonds

```{r The Demand of Diamonds}
library(gridExtra)

plot1 <- ggplot(data=diamonds,aes(x=price))+
    geom_histogram(bins=100)+
    ggtitle('Price')

plot2 <- plot1+
    scale_x_log10()+
    ggtitle('Price (log10)')

grid.arrange(plot1,plot2)
```

***

### Connecting Demand and Price Distributions
Notes: there is bimodality of the price, maybe a "rich vs poor" type of buyer.

***

### Scatterplot Transformation

```{r Scatterplot Transformation}
ggplot(data=diamonds,aes(x=carat,y=price))+
  geom_point()+
  scale_y_log10()
```


### Create a new function to transform the carat variable

```{r cuberoot transformation}
cuberoot_trans = function() trans_new('cuberoot', 
                                      transform =function(x) x^(1/3),
                                      inverse =function(x) x^3)
```

#### Use the cuberoot_trans function
```{r Use cuberoot_trans}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point() + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Overplotting Revisited

```{r Sort and Head Tables}
head(sort(table(diamonds$carat),decreasing=T))
head(sort(table(diamonds$price),decreasing=T))
```


```{r Overplotting Revisited}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point(alpha=1/2,size=3/4,position='jitter') + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Other Qualitative Factors
Notes:

***

### Price vs. Carat and Clarity

Alter the code below.
```{r Price vs. Carat and Clarity}
# install and load the RColorBrewer package
library(RColorBrewer)

ggplot(aes(x = carat, y = price,color=clarity), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```

***

### Clarity and Price
Response: clarity explains a lot of the variance as there is statification across the band due to clarity. holding carat constant, clarity shows a relationship with price.

***

### Price vs. Carat and Cut

Alter the code below.
```{r Price vs. Carat and Cut}
ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Cut', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Cut')
```

***

### Cut and Price
Response: cut seems to have less of a relationship to price than clarity did. Most are ideal cut anyway.

***

### Price vs. Carat and Color

Alter the code below.
```{r Price vs. Carat and Color}
ggplot(aes(x = carat, y = price, color = color), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Color',
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Color')
```

***

### Color and Price
Response: color shows a relationship with price

***

### Linear Models in R
Notes:
We apply the log transformation to the price, and the cube root to the carat.
Response: log(price)~carat^(1/3)

***

### Building the Linear Model
Notes:

```{r Building the Linear Model}
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamonds)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5)
```

Notice how adding cut to our model does not help explain much of the variance
in the price of diamonds. This fits with out exploration earlier.

***

### Model Problems
Video Notes:

Research:
(Take some time to come up with 2-4 problems for the model)
(You should 10-20 min on this)

Response: There are a lot of factors which means it could potentially overfit the data. Also, this may not generalize to new data as diamond price likely fluctuates over time due to inflation and demand etc.

this data is from 2008. inflation and the 2008 recession greatly affected diamond prices from that time. china is also now a factor in diamond prices.
***

### A Bigger, Better Data Set
Notes:

```{r A Bigger, Better Data Set}
# install.package('bitops')
# install.packages('RCurl')
# library('bitops')
# library('RCurl')
# 
# diamondsurl = getBinaryURL("https://raw.github.com/solomonm/diamonds-data/master/BigDiamonds.Rda")
# load(rawConnection(diamondsurl))
```

The code used to obtain the data is available here:
https://github.com/solomonm/diamonds-data

## Building a Model Using the Big Diamonds Data Set
Notes:

```{r Building a Model Using the Big Diamonds Data Set}

diamondsBigSample <- read.csv('diamondsbig.csv')

diamondsBigSample<-subset(diamondsBigSample,price<10000 & cert=='GIA')

m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamondsBigSample)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5)
```


***

## Predictions

Example Diamond from BlueNile:
Round 1.00 Very Good I VS1 $5,601

```{r}
#Be sure you’ve loaded the library memisc and have m5 saved as an object in your workspace.
thisDiamond = data.frame(carat = 1.00, cut = "V.Good",
                         color = "I", clarity="VS1")
modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)

#Need to exp the model estimate because the model is nat log base.
exp(modelEstimate)
thisDiamond
```

Evaluate how well the model predicts the BlueNile diamond's price. Think about the fitted point estimate as well as the 95% CI.

this diamond is decently well fit by the model (5601 actual to 5040 modeled), but the model does have pretty large confidence intervals.

***
```{r}
dat = data.frame(m4$model, m4$residuals)

with(dat, sd(m4.residuals))

with(subset(dat, carat > .9 & carat < 1.1), sd(m4.residuals))

dat$resid <- as.numeric(dat$m4.residuals)
ggplot(aes(y = resid, x = round(carat, 2)), data = dat) +
  geom_line(stat = "summary", fun.y = sd)
```

This shows that the model explains aroun 80ish percent of the variance, and has high variance for the larger diamonds.

## Final Thoughts
Notes:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

