---
title: "Birthdays"
author: "Devin McCormack"
date: "1/24/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Wrangling and EDA on birthdays dataset

```{r}
library(tidyverse)
getwd()
setwd('../lesson4')
list.files()
bday<-read.csv("birthdaysExample.csv")
summary(bday)
```
## Now we have to tidy it up a bit

The data is technically tidy, as it just lists dates. However, we want to be able to probe the data further and could benefit from spliting the data to year, month, day, day of year etc.
```{r}
bday$dates<-parse_date_time(bday$dates,"mdy")
bday<-mutate(bday,year=year(dates),month=month(dates),day=day(dates),doy=yday(dates))
```


```{r}
count(bday,month,sort=TRUE)
ggplot(data=bday,aes(x=month))+
  stat_count()+
  scale_x_continuous(breaks=seq(1,12,1))
```
March contains the most number of birthdays
```{r}
# (count(bday,doy,sort=TRUE))
# max(count(bday,doy,sort=True))
bday %>%
  add_count(doy,sort=TRUE)

ggplot(data=bday,aes(x=day))+
  geom_freqpoly(aes(color=factor(month)),bins=31)+
  labs(x="Day of month",color="Month")
```
There are a bunch of dates with 8 birthdays. and there are no dates with 0 birthdays.
