---
title: "Diamonds EDA"
author: "Devin McCormack"
date: "1/23/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Diamonds EDA

This is an exploration of the diamonds dataset, which is embedded in the ggplot library.

```{r diamonds}
library(tidyverse)
summary(diamonds)
length(diamonds$carat)
length(diamonds)
```

## Quiz 1: 
there are 53940 observations of 10 variables. There are three ordered factors (cut, color, clarity) with the best color being "D"

## Quiz 2:
a price histogram
```{r Price Histogram}
ggplot(data=diamonds,aes(x=price))+
  geom_histogram()

```

## Quiz 3: describe the shape and center of the distribution

```{r price descriptions}
mean(diamonds$price)
median(diamonds$price)
```

The distribution has a long right tail with a mean of $3932.80 and a median of $2401.


## Quiz 4: Diamond counts

How many diamonds cost less than $500?
How many diamonds cost less than $250?
How many diamnods cost $15,000 or more?

```{r price counts}
sum(diamonds$price<500)
sum(diamonds$price<250)
sum(diamonds$price>=15000)
```
# Explore the largest peak in the price histogram you created earlier.

Try limiting the x-axis, altering the bin width, and setting different breaks on the x-axis.

There won’t be a solution video for this question so go to the discussions to share your thoughts and discover what other people find.

You can save images by using the ggsave() command. ggsave() will save the last plot created. For example...
                qplot(x = price, data = diamonds)
                ggsave('priceHistogram.png')

```{r explore the peak}
ggplot(data=diamonds,aes(x=price))+
  geom_histogram(binwidth=50)+
  coord_cartesian(xlim=c(250,2500))
```
## Break out histograms by cut


```{r histogram by cut}
ggplot(data=diamonds,aes(x=price))+
  geom_histogram(binwidth=50)+
  facet_wrap(~cut)
```

## Price by cut

Which has the highest priced diamond?
Which has the lowest priced diamond?
Which has the lowest median price?

```{r}
by(diamonds$price,diamonds$cut,max)
by(diamonds$price,diamonds$cut,min)
by(diamonds$price,diamonds$cut,median)
```
Which has the highest priced diamond? premium
Which has the lowest priced diamond? Premium and Ideal
Which has the lowest median price? Ideal


## Scales for multiple histograms


In the last exercise, we looked at the summary statistics for diamond price by cut. If we look at the output table, the the median and quartiles are reasonably close to each other.


This means the distributions should be somewhat similar,
but the histograms we created don't show that.

The 'Fair' and 'Good' diamonds appear to have 
different distributions compared to the better
cut diamonds. They seem somewhat uniform
on the left with long tails on the right.

Let's look in to this more.

Look up the documentation for facet_wrap in R Studio.
Then, scroll back up and add a parameter to facet_wrap so that
the y-axis in the histograms is not fixed. You want the y-axis to
be different for each histogram.

```{r price per cut}
ggplot(data=diamonds,aes(x=price))+
  geom_histogram(binwidth=50)+
  facet_wrap(~cut,scales='free')
```

# Two variables

## create histogram of price per carat faceted by cut

```{r price per carat by cut}
ggplot(data=diamonds,aes(x=price/carat))+
  geom_histogram()+
  scale_x_log10()+
  facet_wrap(~cut,scales='free')
```

## Investigate the price of diamonds using box plots, numerical summaries, and one of the following categorical variables: cut, clarity, or color.

```{r boxplot of diamond price per carat over colors}
ggplot(data=diamonds,aes(x=color,y=price/carat))+
  geom_boxplot()
by(diamonds$price,diamonds$color,summary)
```


## Inter quartile range IQR

```{r interquartile ranges of color prices}
summary(subset(diamonds,color=='D')$price)
summary(subset(diamonds,color=='J')$price)

IQR(subset(diamonds,color=='D')$price)
IQR(subset(diamonds,color=='J')$price)

```
# Investigate the price per carat of diamonds across the different colors of diamonds using boxplots.


```{r price per carat over colors}
ggplot(data=diamonds,aes(x=price/carat))+
  geom_histogram()+
  scale_x_log10()+
  facet_wrap(~color,scales='free',nrow=3)
```

## carat frequency

```{r carat frequency}
ggplot(data=diamonds,aes(x=carat))+
  geom_freqpoly(binwidth=.01)
```
```{r scatter price v x}
ggplot(data=diamonds,aes(x=x,y=price))+
         geom_point()
```

What is going on? there is some exponential trend

```{r correlations}
with(diamonds,cor.test(x,price))
with(diamonds,cor.test(y,price))
with(diamonds,cor.test(z,price))

```
```{r price v depth}
ggplot(data=diamonds,aes(x=depth,y=price))+
  geom_point(alpha=1/100)+
  scale_x_continuous(breaks=seq(2,80,2))

```

typical depth range is between 58 and 64.

```{r corr depth and price}
with(diamonds,cor.test(depth,price))
```

```{r scatter price v carat}
ggplot(data=subset(diamonds,price<quantile(price,.99) & carat<quantile(carat,.99)),
       aes(x=carat,y=price))+
  geom_point()
```

```{r scatter price vs rough volume}
diamonds%>%
  mutate(volume=x*y*z)->diamonds
ggplot(data=diamonds,aes(x=volume,y=price))+
  geom_point()
```

How many diamonds have 0 volume??
```{r diamond volume}
sum(diamonds$volume==0)
```
## what is the correlation of price and volume, excluding diamonds with volume=0 or >=800

```{r correlation price and volume}
with(subset(diamonds,volume>0 & volume<=800),cor.test(price,volume))
```
## price vs volume
We encourage you to think about this next question and to post your thoughts in the discussion section.

Do you think this would be a useful model to estimate the price of diamonds? Why or why not?

```{r fit o}
ggplot(data=subset(diamonds,volume>0 & volume<=800),
       aes(x=volume,y=price))+
  geom_point(alpha=1/50)+
  geom_smooth()
```

This does not seem like a great fit because it shows that at some point, diamonds get cheaper per volume, which is very false.


Use the function dplyr package
to create a new data frame containing
info on diamonds by clarity.

Name the data frame diamondsByClarity


```{r}
diamondsByClarity<-diamonds%>%
  group_by(clarity)%>%
  summarise(mean_price=mean(price),
            median_price=median(price),
            min_price=min(price),
            max_price=max(price),
            n=n())
head(diamondsByClarity)
```
```{r two bar plots}
library(gridExtra)

diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

diamonds_mp_by_clarity
dclar<-ggplot(data=diamonds_mp_by_clarity,aes(x=clarity,y=mean_price))+
  geom_col()

dcol<-ggplot(data=diamonds_mp_by_color,aes(x=color,y=mean_price))+
  geom_col()

grid.arrange(dclar,dcol,ncol=1)
```

# Multiple variables

## Create a histogram of diamond prices. 
Facet the histogram by diamond color and use cut to color the histogram bars.

```{r diamond prices by color filled by cut}
ggplot(data=diamonds,aes(x=price,fill=cut))+
  geom_histogram(bins=50)+
  scale_x_log10()+
  facet_wrap(~color)+
  scale_fill_brewer(type='qual')
  
```

## Create a scatterplot of diamond price vs. table and color the points by the cut of the diamond.
```{r price vs table colored by cut}
ggplot(data=diamonds,aes(y=price,x=table,color=cut))+
  geom_point()+
  scale_color_brewer(type='qual')
```

What is the typical table range for diamonds of ideal cut?

53-57

What is the typical table range for diamonds of premium cut?

58-62

## Create a scatterplot of diamond price vs. volume (x * y * z) and color the points by the clarity of diamonds. 
Use scale on the y-axis to take the log10 of price. 
You should also omit the top 1% of diamond volumes from the plot.

```{r price vs volume color by clarity}
ggplot(data=subset(diamonds,volume<quantile(volume,.99)&
                     volume>0),
       aes(x=volume,y=price,color=clarity))+
  geom_point()+
  scale_y_log10()+
  scale_color_brewer(type='div')
  
```

## Create a scatter plot of the price/carat ratio of diamonds. The variable x should be assigned to cut. The points should be colored by diamond color, and the plot should be faceted by clarity.

```{r}
ggplot(data=diamonds,aes(x=cut,y=price/carat,color=color))+
  geom_jitter(size=1)+
  facet_wrap(~clarity)+
  scale_color_brewer(type='div')
```


# Linear Regression models

Scatterplot review

```{r}
ggplot(data=subset(diamonds,carat<quantile(carat,.99)&price<quantile(price,.99)),aes(x=carat,y=price))+
  geom_point(fill=I('#F79420'),color=I('black'),shape=21)
```




