Red Wine Quality by Devin McCormack
========================================================


```{r echo=FALSE, message=FALSE, warning=FALSE, "packages"}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.
library(tidyverse)
library(gridExtra)
```

```{r echo=FALSE, "Load the Data"}
# Load the Data
wine<-read.csv('wineQualityReds.csv')


```

This dataset tabulates the physicochemical properties, as well
as subjective quality (1-10) of 1599 red wines  from: 
P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis.
Modeling wine preferences by data mining from physicochemical properties.
In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.

The specific wine variety tested is a Portuguese wine called "Vinho Verde".
There are 11 attributes of these wines mearued with physicochemical tests,
most measurements in grams (or milligrams) per liter. Exceptions are pH, which 
is a scale, and alcohol, which is given in percentage by volume.


These 11 attributes are measured in an attempt to model a "quality" score.
Here quality is a median of at least 3 evaluations made by wine experts.
Wine Quality is a complete sensory score influenced by the wine tasting process,
which is generally laid out in the five "S" steps:

1. See - the visual color and clarity of the wine is judged. This is not
directly related to any of our measured components.

2. Swirl - the density, related to alcohol and residual sugar is noted.

3. Sniff - the "nose" of the wine is most important. From here a judge can 
assess the fermentation (acetic acid), alcohol, preservatives (sulfur dioxide),
and exotic aromas (citric acid and others) of a wine.

4. Sip - tasting the wine allows a judge to assess sweetness (residual sugar),
tartness (acids and pH), salts (chlorides), and alcohol content. Additionally,
various elements of smells change with temperature. Interestingly, a key 
component of taste is usually bitterness or tannin level, but the data does
not reflect that this was physicochemically measured.

5. Savor - the aftertaste of the wine more completely assesses alcohol and 
bitterness, as well as less volatile aromatics.

Now into data cleaning!

## Data cleaning

```{r echo=FALSE, "summary"}

summary(wine)

```

Based on the summary output there are two things I might want to change: 
first dropping the unnecessary X column, and changing quality to a factor. 

One thing worth noting right now is that the quality range isn't 1-10. 
There are no absolutely terrible or perfect red wines in this list.

At this point it is important to note the units of these measurements. 
All input variables besides pH, density, sulfur dioxides, and alcohol are 
in g/dm^3, aka grams per liter. Density is in g/cm^3, or grams per milliliter. 
This can be converted to g/dm^3 by multiplying by 1000. Alcohol is given in 
percent by volume. Knowing that the density of pure ethanol is 789 g/dm^3, we 
can also convert percent into g/dm^3. Both sulfur dioxide measurements are in 
mg/dm^3, so we can divide by 1000 to get it into grams.

Finally, pH is a base 10 scale of acidity. The different acids measured 
(i.e. tartaric, acetic, citric) have different acidity levels based on 
concentrations and other components in the water. The sum of these components 
should be negatively correlated with pH.

```{r "Explore and clean data"}

wine<-select(wine,-X)
wine$quality<-as.factor(wine$quality)

wine%>%
  mutate(density=density*1000,
         alcoholgperL=alcohol*7.89,
         free.sulfur.dioxide=free.sulfur.dioxide/1000,
         total.sulfur.dioxide=total.sulfur.dioxide/1000)->wine

head(wine)
```

# Univariate Plots Section

## Quality Distribution

First a quick look at what we are trying to model, the quality score of the wine.
<!-- > **Tip**: In this section, you should perform some preliminary exploration of -->
<!-- your dataset. Run some summaries of the data and create univariate plots to -->
<!-- understand the structure of the individual variables in your dataset. Don't -->
<!-- forget to add a comment after each plot or closely-related group of plots! -->
<!-- There should be multiple code chunks and text sections; the first one below is -->
<!-- just to help you get started. -->

```{r echo=FALSE, "Univariate_Plots, quality distribution"}
ggplot(data=wine,aes(x=quality))+
  geom_bar()+
  scale_y_log10()

```

Here we see the distribution of the quality ratings. Once again, this shows that
there is really a constricted range of values; all wines are rated 3-8, not 1-10. 
Note that the y axis of data is also plotted on log10 scale. This is because
there is an order of magnitude less wines at the extremes than in the middle.
It is rare for wines to be much better or worse than average.

## Acids

Acids are important to wine, and the type and amount of acidity affects the
smell, sip, and savor steps of winetasting.

```{r echo=FALSE, "Univariate_Plots, total acids"}
p.fixed<-ggplot(data=wine,aes(x=fixed.acidity,))+
  geom_histogram(binwidth = .5,aes(color=I('gray'),fill=I('mediumpurple')))
p.volatile<-ggplot(data=wine,aes(x=volatile.acidity,))+
  geom_histogram(binwidth = .05,aes(color=I('gray'),fill=I('tomato')))
p.citric<-ggplot(data=wine,aes(x=citric.acid,))+
  geom_histogram(binwidth = .025,aes(color=I('gray'),fill=I('limegreen')))
p.pH<-ggplot(data=wine,aes(x=pH,))+
  geom_histogram(binwidth = .05,aes(color=I('gray'),fill=I('aquamarine')))

grid.arrange(p.fixed,p.volatile,p.citric,p.pH,ncol=1)

```

Noteworthy are that all the wines have some amount of fixed acidity and volatile
acidity, but not all wines have citric acid. Also noteworthy is the different
scales of these graphs. Fixed acid is a measure of tartaric acid, which is a
product of grapes - it makes sense that wine from grapes has a lot of it. 
Volatile acid is acetic acid, a component of vinegar, which is a byproduct of 
fermentation. Since wine is fermented, we expect some level of acetic acid, but
no one wants particularly vinegary wine. Citric acid is related to citrus fruit,
and is rarer to find in wine. However it is noted to be a pleasant addition.
pH shows a fairly gaussian distribution, indicating that the total acidity is
fairly normal, and centered around a range that similar to that of grapes.
Later it will be worth looking at how correlated these variables are.


## Density

Density is a major component of both swirling and sipping, and therefore may be
a big factor in quality.

```{r echo=FALSE, "Univariate_Plots, density"}
ggplot(data=wine,aes(x=density))+
geom_histogram(binwidth=1,color=I('gray'),fill=I('cornflowerblue'))

```

The peak density is below 1000, meaning that in general, the wine is less dense 
that pure water. This makes sense, since alcohol has lower density than water
and it generally makes up ~10% of the volume.

## Alcohol content

Alcohol affects the swirl, sip, and savor components. Alcohol also has effects
on physiological conditions, and could influence scoring beyond the 5 "s" steps.

```{r echo=FALSE, "Univariate_Plots, Alcohol content"}
ggplot(data=wine,aes(x=alcohol))+
geom_histogram(binwidth=.5,color=I('gray'),fill=I('cornflowerblue'))

```

This data shows that most of the wines have between 9 and 11% alcohol by volume,
and there seems to be a positive skew.

## Residual sugar content

Sugar affects see, sip, and savor.

```{r echo=FALSE, "Univariate_Plots, Sugar content"}

sugarhist<-ggplot(data=wine,aes(x=residual.sugar))+
geom_histogram(binwidth=.5,color=I('gray'),fill=I('cornflowerblue'))

sugarbox<-ggplot(data=wine,aes(x=1,y=residual.sugar))+
  geom_jitter(alpha=.3,color='cornflowerblue')+
  geom_boxplot(alpha=.5,fill=I('cornflowerblue'))

grid.arrange(sugarhist,sugarbox,ncol=2)

```

Most wines fall below 4 g/L, indicating that this wine type is likely a dry wine.
There is a long right tail, but all are significantly less than the threshold
for a sweet wine, 45g/L. The fermentation process converts sugar to alcohol, so
it would be reasonable that, if starting sugar was constant, that alcohol and 
residual sugar would have a negative correlation.

A boxplot explores these outliers even more. Generally, wines are called "dry"
if they have less than 10 g of sugar, making it likely that it is hard to
distinquish sugar levels much at these levels. An expert likey will be able to 
taste that several of these outliers are "off-dry", but unless there is a clear
trend in the bivariate plots, it likely isn't important for these wines.

## Sulfur Dioxide (preservatives)

Sulfur Dioxide (SO2) is a preservative of wine, and can affect the sniff and sip. 
Too little preservatives can lead to spoiled wine, however it is unlikely that 
any of the tested wines were spoiled. Too much can give a sulfur taste, which
is almost universally a bad taste for wine.


```{r echo=FALSE, fig.width=8.5, "Univariate_Plots, SO2 content"}

p.free<-ggplot(data=wine,aes(x=free.sulfur.dioxide,))+
  geom_histogram(bins=30,aes(color=I('gray'),fill=I('lightgoldenrod')))
p.total<-ggplot(data=wine,aes(x=total.sulfur.dioxide,))+
  geom_histogram(bins=50,aes(color=I('gray'),fill=I('goldenrod')))
p.sulphates<-ggplot(data=wine,aes(x=sulphates,))+
  geom_histogram(bins=50,aes(color=I('gray'),fill=I('darkgoldenrod')))

grid.arrange(p.free,p.total,p.sulphates,ncol=3)

```

We see all three graphs have a right skew. All of these should be correlated, 
as total SO2 includes free SO2, and increased sulphates increases SO2.

## Chlorides (salt)

It is unclear how much this affects quality, as I've never had a "salty" wine, 
but it can potentially manifest as a mineral, metallic, or savory taste, and 
influence texture - both components of sip.

```{r echo=FALSE, "Univariate_Plots, Salt content"}
salthist<-ggplot(data=wine,aes(x=chlorides))+
geom_histogram(binwidth=.01,color=I('gray'),fill=I('cornflowerblue'))

saltbox<-ggplot(data=wine,aes(x=1,y=chlorides))+
  geom_jitter(alpha=.3,color='cornflowerblue')+
  geom_boxplot(alpha=.5,fill=I('cornflowerblue'))

grid.arrange(salthist,saltbox,ncol=2)


```

There appears to be a normal distribution near .1g/L, with a very long right
tail. It's hard to say if such low levels of chlorides has any appreciable
affect on taste.

<!-- > **Tip**: Make sure that you leave a blank line between the start / end of -->
<!-- each code block and the end / start of your Markdown text so that it is -->
<!-- formatted nicely in the knitted text. Note as well that text on consecutive -->
<!-- lines is treated as a single space. Make sure you have a blank line between -->
<!-- your paragraphs so that they too are formatted for easy readability. -->

# Univariate Analysis

Generally, we see that features are normal with a right skew, with density being
an exception with no discernable skew. The main feature, quality, shows that 
exceptional wines (both bad and good) are very uncommon. This can be due to
the aggregation of the quality score (median of 3 or more tastes),
psychological mean reversion due to contrast of qualitative measurements over
time, or just a overall small difference between wines of this variety.

### What is the structure of your dataset?

Data is structured where each row is a single bottle of Vihno Verde wine. There
is no information about grape variety, vintage, winery, brand, or price. 
Each wine has physicochemical measurements as well as a median aggregated 
measure of quality from 3+ tastings by wine experts.

### What is/are the main feature(s) of interest in your dataset?

The main feature of interest is quality, and we are using the rest of the
features to try and model quality. 

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

The features of most interest to me are features that predominately affect 
sniff, or multiple of the five "S" steps of tasting. These include alcohol, 
sugar, acids, and sulfur dioxide.

### Did you create any new variables from existing variables in the dataset?

I created a new alcohol variable to turn percent volume into g/L, but it is 
still linearly related to alcohol, so it is not neccessarily a new variable in
terms of linear regression.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

I converted features, when possible, to all have the same units of g/L. Due to 
a right skew, many of the features can be scaled with log10, however I didn't
convert the raw numbers to log10 scale at this time.

# Bivariate Plots Section

<!-- > **Tip**: Based on what you saw in the univariate plots, what relationships -->
<!-- between variables might be interesting to look at in this section? Don't limit -->
<!-- yourself to relationships between a main output feature and one of the -->
<!-- supporting variables. Try to look at relationships between supporting variables -->
<!-- as well. -->

## ggpairs analysis

First I can look at all the comparisions quickly with ggpairs, and then look
closer with bigger plots.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8.5, fig.height=8.5,"ggpairs analysis"}
library(GGally)
library(scales)
library(memisc)

ggpairs(wine,
  lower = list(continuous = wrap("points", shape = I('.'))),
  upper = list(combo = wrap("box", outlier.shape = I('.'))),
  columnLabels = c("tartaric","acetic","citric","sugar","salt","free_SO2",
                   "bound_SO2","density","pH","sulphates","alcohol",
                   "quality","g alcohol"))+
  theme(axis.ticks=element_blank(),axis.text.x = element_blank(),
        axis.text.y = element_blank())
```

Looking at the ggpairs (.pdf in the git folder), we can see a few trends. 
In the scatter plots, we can see a predictable correlation between various 
acidities and the pH. Also interesting to see trends between fixed acidity and 
density, and alcohol and density. This makes sense since those are the largest 
values in g/L, meaning they add the most mass to the liquid. Looking at the 
sugar boxplots, there doesn't seem to be a trend with sugar content, and the
outliers seem to be all pretty mediocre grades. This confirms my thought that
sugar likely is an afterthought in this type of wine.

However, the most question central to this EDA is: what attributes affect the 
overall quality score of the wine? Looking at the boxplots for quality, we see 
there may be a downward trend in volatile acidity that affects quality - where 
more acetic acid in the wine may affect quality - likely by affecting the smell
of the wine. Conversely, there looks to be an upwards trend with citric acid. 
Citric acid is noted to give the wine a "fresh" smell, which may boost quality. 
One more that stands outis that higher quality wines seems to have higher 
alcohol content - maybe a small change in the tipsyness of judges affects how 
they grade for quality?

## Acids vs pH

```{r echo=FALSE, "Bivariate_Plots total acids vs pH"}
ggplot(data=wine,aes(x=fixed.acidity+volatile.acidity+citric.acid,
                     y=pH))+
  geom_point(alpha=.5,color=I('aquamarine3'))+
  geom_smooth(method='lm')

with(wine,cor.test(fixed.acidity+volatile.acidity+citric.acid,pH))
```


```{r echo=FALSE, "Bivariate_Plots fixed acid vs pH"}
ggplot(data=wine,aes(x=fixed.acidity,y=pH))+
  geom_point(alpha=.5,color=I('mediumpurple'))+
  geom_smooth(method='lm')

with(wine,cor.test(fixed.acidity,pH))
```

The sum of the acid components is highly correlated with the pH, which makes
sense, and makes pH a bit redundant. Looking at boxplots in ggpairs, the type
of acid is more important than overall acidity, within the ranges we see in 
this dataset. It is noteworthy that it seems that fixed acidity is the driving
factor in the sum, so if we seek to drop pH, we should include fixed acidity.

## Fixed acidity vs density

looking at the correlation between fixed acidity, the largest contribution of
mass besides alcohol to the wine.

```{r echo=FALSE, "Bivariate_Plots fixed acidity vs density"}
ggplot(data=wine,aes(x=fixed.acidity,y=density))+
  geom_point(alpha=.5,color=I('mediumpurple'))+
  geom_smooth(method='lm')

with(wine,cor.test(fixed.acidity,density))

```

We see a lot of correlation between these two, which makes sense since tartaric
acid is the largest dissolved contribution to mass besides alcohol.

## Alcohol vs density

Alcohol should also have a correlation with density, but it will be negative,
since alcohol has less density than pure water.

```{r echo=FALSE, "Bivariate_Plots alcohol vs density"}
ggplot(data=wine,aes(x=alcohol,y=density))+
  geom_point(alpha=.5,color=I('cornflowerblue'))+
  geom_abline(slope=-2.11,intercept=1000,linetype=2)+
  scale_y_continuous(limits=c(970,1010))

with(wine,cor.test(alcohol,density))

```


Here we see that alcohol and density are negatively correlated, but there are 
other components that affect density as well. A mix of ethanol and pure water
would follow the density curve of density=-2.11*(alcohol% by vol)+1000, but the
points are all well above these values, indicating that there are heavier than
water components in wine, like tartaric acid, sugar, etc.


## Total vs Free sulfur dioxide

total sulfur dioxide (SO2) includes free SO2 in the measurment. It may be
interesting to see what the ratio, of free SO2 to bound SO2 is in wines.

```{r echo=FALSE, fig.width=8.5,"Bivariate_Plots total vs free sulfur dioxide"}
ggplot(data=wine,aes(x=total.sulfur.dioxide-free.sulfur.dioxide,
                     y=free.sulfur.dioxide))+
  geom_point(alpha=.5,color=I('goldenrod'))+
  geom_abline(slope=1,linetype=2)+
  geom_abline(slope=1/8)+
  geom_abline(slope=8)+
  coord_fixed()+
  xlab('bound.sulfur.dioxide')

with(wine,cor.test(total.sulfur.dioxide-free.sulfur.dioxide,free.sulfur.dioxide))

```

Here, plotted with a dashed unity line (which would indicate 50% free SO2), 
and 2 solid lines indicating rough bounds. The boundaries of the data seem to
be factors of 8, meaning that at most, there is ~8 times as much free SO2 than
bound, and vice versa. the equilibrium between the two forms is likely highly 
dependent on solution contents (and temperature), but it is not likely that 
there are equilibrium points that create more than 8x as much of one form than 
the other. We see that most wines have less than 50% of their SO2 in free form.
There doesn't seem to be an obvious trend, especially if the two bound SO2
points are discounted.

## total sulfur dioxide and sulphates

We might expect these to be correlated, since sulphates are often added to wine
as to boost SO2, however we might also see that they are uncorrelated because
adding sulphates may not be the only way to maintain SO2 levels in the wine.

```{r echo=FALSE, "Bivariate_Plots Sulphates vs total Sulfur dioxide"}
ggplot(data=subset(wine,total.sulfur.dioxide<.2),
       aes(x=total.sulfur.dioxide,y=sulphates))+
  geom_point(alpha=.5,color=I('goldenrod'))+
  geom_smooth(method='lm') #removed 2 oulier points

with(wine,cor.test(total.sulfur.dioxide,sulphates))

```

We see very little correlation between to two, and this might indicate that
sulphates may not be the only way to maintain SO2, or that sulphates are only
added to correct low SO2 levels, and not added uniformly.


## Quality box plots

Importantly, we need to look at what features seem to vary with quality.

```{r echo=FALSE, "quality vs citric and volatile acids"}

q.cacid<-ggplot(data=wine,aes(x=quality,y=citric.acid))+
  geom_jitter(alpha=.3,color='limegreen')+
  geom_boxplot(alpha=.5,fill='limegreen')+
  stat_summary(fun.y="mean",geom="point",color="red",shape=8,size=4)
q.vacid<-ggplot(data=wine,aes(x=quality,y=volatile.acidity))+
  geom_jitter(alpha=.3,color='tomato')+
  geom_boxplot(alpha=.5,fill='tomato')+
  stat_summary(fun.y="mean",geom="point",color="green",shape=8,size=4)

grid.arrange(q.cacid,q.vacid,ncol=1)

```

Boxplots are plotted over a jittered scatter of the data, with means represented
with a contrasting star point.
One trend that stands out in the ggpairs plots is that higher rated wines tend 
to have less volatile acids, and more citric acids. Both of these components 
can affect the nose of a wine, with volatile acids presenting as vinegar, and 
citric acid presenting as freshness. I think that maybe the ratio of acids is 
more important than the total amounts. Noteworthy is how far the 3 quality 
median value of citric acid is below the mean. This means that there are a few 
big outliers, but most are near zero.

```{r echo=FALSE, "quality vs alcohol and sulfates"}

q.alcohol<-ggplot(data=wine,aes(x=quality,y=alcohol))+
  geom_jitter(alpha=.3,color='cornflowerblue')+
  geom_boxplot(alpha=.5,fill='cornflowerblue')+
  stat_summary(fun.y="mean",geom="point",color="red",shape=8,size=4)
q.sulfur<-ggplot(data=wine,aes(x=quality,y=sulphates))+
  geom_jitter(alpha=.3,color='goldenrod')+
  geom_boxplot(alpha=.5,fill='goldenrod')+
  stat_summary(fun.y="mean",geom="point",color="red",shape=8,size=4)

grid.arrange(q.alcohol,q.sulfur,ncol=1)

```

Alcohol shows a pretty inticing trend, with higher quality wines having higher
alcohol content. At first glance, we may attribute this to some fractional
higher level of happiness that the expert may feel after drinking more alcohol,
but it is important to also consider that changing the alcohol concentration in
the solution may greatly affect the soluability of volatile compounds that 
affect the smell and taste of the wine.

Sulphates show a mild trend, maybe indicating that better wines use more SO2 to
prevent even the slightly amount of spoilage or improper fermentation, but it is
hard to say it is conclusive since there is such a small difference between 
levels, and there are a lot of outliers for middling wines.

# Bivariate Analysis


### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

We see that quality is positively related to alcohol, citric acid, and 
sulphates, and seems to be negatively related to volatile acidity.

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

We see that the sum of acids is highly correlated with the pH, which means that
pH may be a redundant measurement. We also see high correlations between
alcohol and density, and fixed acid and density.

### What was the strongest relationship you found?

the strongest relationship I found was between sum total of acids and the pH, 
with a correlation coefficient of -0.68. more mass of acids means lower pH.


# Multivariate Plots Section

## Volatile vs citric acid colored by quality

Above we saw that volatile and citric acid seemed to have opposite relationships
with quality. Here we plot them against each other, colored by quality to see if 
any grouping falls out in the data.

```{r echo=FALSE, "Multivariate_Plots, citric acid vs volatile acid, quality"}
ggplot(data=wine,aes(color=quality,x=citric.acid,y=volatile.acidity))+
  geom_point(alpha=.5)+
  geom_smooth(method='lm',se=FALSE,size=.5)+
  geom_abline(slope=1,linetype=2)+
  scale_color_brewer(palette='RdYlGn', 
                     guide = guide_legend(title = 'Quality', reverse = T, 
                                          override.aes = list(alpha = 1, 
                                                              size = 2)))

```

Here with the dashed unity line showing equal parts volatile and citric acid.
Interestingly, we see some grouping of better quality wines looking to have 
more citric acid than volatile acid. Most interestingly is that the bad 
wines seem to be grouped along the upper y axis, indicating they have little to
no citric acid, and higher levels of volatile acid than most other wines.

## Sulphates vs volatile acids

Considering how bunched the bad wines are, maybe a lot of volatile acids
indicates some secondary fermentation that is usually stopped by sulphate
preservatives.

```{r echo=FALSE, "Multivariate_Plots, sulphates vs volatile acid, quality"}
ggplot(data=wine,aes(color=quality,x=sulphates,y=volatile.acidity))+
  geom_point(alpha=.5)+
  geom_smooth(method='lm',se=FALSE,size=.5)+
  scale_color_brewer(palette='RdYlGn', 
                     guide = guide_legend(title = 'Quality', reverse = T, 
                                          override.aes = list(alpha = 1, 
                                                              size = 2)))

```

Interestingly we see some bunching again. worse wines usually have high volatile
acids as well as low sulphates. Mayber there is some creedence to the idea that
worse wines are not as carefully maintained, and have secondary fermentations
that give off flavors.

## Alcohol and citric acid

Alcohol and citric acid looked to both be positively related with quality.

```{r echo=FALSE, "Multivariate_Plots, citric acid vs alcohol and quality"}
ggplot(data=wine,aes(x=citric.acid,y=alcohol,color=quality))+
  geom_point(alpha=.5)+
  geom_smooth(method='lm',se=FALSE,size=.5)+
  scale_color_brewer(palette='RdYlGn',
                     guide = guide_legend(title = 'Quality', reverse = T,
                                          override.aes = list(alpha = 1,
                                                              size = 2)))
```

We see some bunching of green in the top right, indicating high alcohol+ high
citric acid usually means a better wine.


## Modeling the data

I use ordinal regression, as we have a multiclass variable we are trying to
describe, and the classes are ordered. This is done with the polr function.

I include  all three acids, alcohol, and sulphates in the model.

```{r echo=FALSE, "Multivariate_Plots, modeling with ordinal regression"}

m1 <- polr(quality ~ citric.acid, data = wine,Hess=TRUE)
m2 <- update(m1, ~ . + fixed.acidity)
m3 <- update(m2, ~ . + volatile.acidity)
m4 <- update(m3, ~ . + alcohol)
m5 <- update(m4, ~ . + sulphates)
mtable(m1, m2, m3, m4, m5)

```


```{r echo=FALSE, "Multivariate_Plots, Confusion matrix"}

modelEstimate = predict(m5, newdata = wine)


ftable(modelEstimate,wine$quality)
```

This model has high misclassifications, but seems to trend in the right 
direction. There are no gross misclassifications, save one "5" being judged as
the sole "8".


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

One of the most visually inticing relationships that I plotted was citric acid
vs volatile acid, colored by quality. It seemed that broadly, if the wine had
more citric acid than volatile acids, it was good, and high volatile acids in
general lead to worse wines.

A clear case of strengthening each other can be made for citric acid and
alchol. It seems that the higher alcohol content wines with higher citric acid
levels were generally better wines.


### Were there any interesting or surprising interactions between features?

It is interesting, but not surprising, that high acetic (volatile) acid wines
were generally worse. Our sense of smell is very sensitive to the vinegar
smell, and it indicates the wine potentially has secondary fermentations that
are usually undesireable. This is backed up somewhat by the sulphates vs
volatile acids. Potentially, there are other smells and flavors that are not
measured that result from similar fermentations that increase volatile acids.

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

Interestingly, the model seems to compress the data even further, misclassifying
most wines in the 5/6 categories. It seems to not gravely mistake bad wines for
good, or vice versa, so it is an intriguing starting point. The downfall of this
model likely is the small sample size, especially for especially good or bad
wines. When there are only 18 out of 1599 wines that are graded as an "8", it is
hard to generalize. Interestingly, the only wine that the model classifies as an
8 is actually a 5.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

ggplot(data=wine,aes(color=quality,x=citric.acid,y=volatile.acidity))+
  geom_point(alpha=.5)+
  geom_abline(slope=1,linetype=2)+
  geom_smooth(method='lm',se=FALSE,size=.5)+
  scale_color_brewer(palette='RdYlGn', 
                     guide = guide_legend(title = 'Quality', reverse = T, 
                                          override.aes = list(alpha = 1, 
                                                              size = 2)))+
  xlab("Citric Acid Concentration g/L")+
  ylab("Volatile (Acetic) Acid Concentration g/L")+
  ggtitle('Wine Quality: Good Wines have more Citric Acid than Acetic Acid')

```

### Description One

Here we see that higher quality wines tend to have greater than a 1:1 ratio of
citric acid to volatile acid. This is most evident when looking at poor wines,
which tend to have no citric acid and high amounts of volatile acid. The linear
fit for each quality type shows that volatile acid seems to be the important
stratifier for quality. This plot shows the importance of these two acids to
wine quality.

### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(data=wine,aes(color=quality,x=sulphates,y=volatile.acidity))+
  geom_point(alpha=.5)+
  geom_smooth(method='lm',se=FALSE,size=.5)+
  scale_color_brewer(palette='RdYlGn', 
                     guide = guide_legend(title = 'Quality', reverse = T, 
                                          override.aes = list(alpha = 1, 
                                                              size = 2)))+
  xlab("Added Potassium Sulphate g/L")+
  ylab("Volatile (Acetic) Acid Concentration g/L")+
  ggtitle('Wine Quality: Volatile Acid vs. Sulphate concentration')

```

### Description Two

Here we can see that bad wines have high acetic acid, and low sulphates. We also
see that outliers with very high potassium sulphates tend to be poor if volatile
acids are high. We can justify this by understanding that potassium sulphate is
added to wine as a preservative, and if added in too low a quantity, or too
late in the process, can cause secondary fermentations to occur. Great 
winemakers likely can add the minimal amount of sulphates to wine to prevent
spoilage, while worse winemakers may over-react and add too much.


### Plot Three
```{r echo=FALSE, Plot_Three}
ggplot(data=wine,aes(x=citric.acid,y=alcohol,color=quality))+
  geom_point(alpha=.5)+
  geom_smooth(method='lm',se=FALSE,size=.5)+
  scale_color_brewer(palette='RdYlGn',
                     guide = guide_legend(title = 'Quality', reverse = T,
                                          override.aes = list(alpha = 1,
                                                              size = 2)))+
  xlab("Citric Acid Concentration g/L")+
  ylab("Alcohol % (vol/vol)")+
  ggtitle('Wine Quality: The more Alcohol and Citric Acid, the better')
```

### Description Three

Higher quality wines tend to have higher than average alcohol and citric acid
content. This graph is important because it shows pretty nice stratification
with the linear models. The lower quality wines have lower alcohol than higher
quality wines. The negative slope in the highest quality group indicates a
trend that lower alcohol wines can still be highly rated if they have higher
citric acid content.

------

# Reflection

This was a fun dataset, that lead to some interesting insights into the grading
of wines. Both wine making and tasting are such complex processes, that even
small differences in physicochemical measurements can compound into large
differences in the final quality of the product. Some trends, such as
having high acetic acid reducing the quality of wine, seem to agree with 
conventional wisdom of even novice winetasters. Others, such as positive
association with alcohol content and quality, resonate with people who think
that winetasting may all be psychosomatic - the percieved ability to contrast 
and describe wines is much more important than any actual small physicochemical
difference makes in the satisfaction and quality of wine.

This dataset, however, suffered from being small on multiple fronts. The quality
metric was compressed and did not span the entire range of values. Modeling the
data was hamstrung by the fact that a vast majority of wines were "okay",
very few were "good" or "bad", and none were "exceptional" or "terrible".

I feel that the measurements lacked at least one crucial component in wine:
tannins. Additionally, after researching Vihno Verde, I beleive that lactic
acid and CO2 are two components that would be crucial to measure. And finally,
it would have been a much more multifaceted and interesting analysis if there 
were some more qualitative measures, like categorization of tasting notes.
Is there a threshold of acetic acid before a taster notes vinegar? How finely
can an expert distinguish alcohol content? What acid balance is associated with
different fruit comparisons?

However, the constraint of the dataset also made it easier to approach. 
Measuring a single type of wine, irrespective of grape varietal, growth region,
vintage, etc. allows for a more focused approach without an explosion of 
factors. 

Future work with this type of data, I think, is very much dependent on gathering
more data! Describing qualitative feelings with quantitative measures is a
major end goal of analytics, and wine tasting is a endless source of this data,
it just needs to be collected.












