Lesson 5
========================================================

### Multivariate Data
Notes: scatterplots are the simpliest form of multivarite visualization

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:
```{r}

library(tidyverse)
pf<-read.csv('pseudo_facebook.tsv',sep='\t')

ggplot(data=pf,aes(x=age,y=friend_count))+
  geom_point()

```

```{r Third Qualitative Variable}

ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot()+
  stat_summary(fun.y=mean,geom="point",shape=4)

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender)))+
  geom_line(aes(color=gender),stat='summary',fun.y=median)
```


Write code to create a new data frame,
called 'pf.fc_by_age_gender', that contains
information on each age AND gender group.

The data frame should contain the following variables:

mean_friend_count,
median_friend_count,
n (the number of users in each age and gender grouping)

```{r}
pf.fc_by_age_gender<-pf%>%
  filter(!is.na(gender))%>%
  group_by(age,gender)%>%
  summarise(mean_friend_count=mean(friend_count),
            median_friend_count=median(friend_count),
            n=n())
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(data=pf.fc_by_age_gender,
       aes(x=age,y=median_friend_count,color=gender))+
  geom_line()
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes: long formats has many repeats, it is tidy. wide format is cleaner to look at without SQL type data packaging.

***

### Reshaping Data
Notes:

```{r}
# install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender.wide<-dcast(pf.fc_by_age_gender,
                                age~gender,
                                value.var='median_friend_count')
head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
library(tidyverse)
ggplot(data=pf.fc_by_age_gender.wide,aes(x=age,y=female/male))+
  geom_line()+
  geom_hline(linetype=2,alpha=.3,yintercept=1)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf<-mutate(pf,year_joined=2014-ceiling((tenure/365)))
# below is garbage version answer they want
# pf$year_joined<-with(pf,2014-ceiling((tenure/365)))

```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
summary(pf$year_joined)
table(pf$year_joined)

pf$year_joined.bucket<-with(pf,cut(year_joined,breaks=c(2004,2009,2011,2012,2014)))
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket)))+
  geom_line(aes(color=year_joined.bucket),stat='summary',fun.y=median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket)))+
  geom_line(aes(color=year_joined.bucket),stat='summary',fun.y=mean)+
  geom_line(stat='summary',fun.y=mean,linetype=2)
```

***

### Friending Rate
Notes: friends per day of tenure

```{r Friending Rate}
with(subset(pf,tenure>0),summary(friend_count/tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

median friend count is .2205, and max is 417

Create a line graph of mean of friendships_initiated per day (of tenure) vs. tenure colored by year_joined.bucket.
```{r Friendships Initiated}
ggplot(data=subset(pf,tenure>0),
       aes(x=tenure,y=friendships_initiated/tenure))+
  geom_line(stat='summary',fun.y=mean,aes(color=year_joined.bucket))
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)


```
```{r}
## use geom_smooth for same effect
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))
```

***

### Sean's NFL Fan Sentiment Study
Notes:the correct average frame is important to use. for NFL, a 7 day moving average makes sense since they usually have 1 game a week.

***

### Introducing the Yogurt Data Set
Notes: there is a history of purchase data over time. lets look at household purchases of 5 flavors of yogurt. many rows per household, one per purchase.

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo<-read.csv('yogurt.csv')

yo$id<-factor(yo$id)
str(yo)

ggplot(data=yo,aes(x=price))+
  geom_histogram(binwidth=10)
```

***

### Number of Purchases
Notes:Create a new variable called all.purchases, which gives the total counts of yogurt for each observation or household.

```{r Number of Purchases}
summary(yo)
length(yo)
table(yo$price)

yo<-transform(yo,all.purchases=strawberry+blueberry+pina.colada+plain+mixed.berry)

ggplot(data=yo,aes(x=all.purchases))+
  geom_histogram()
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(data=yo,aes(x=time,y=price))+
  geom_jitter(alpha=1/20)
```

***

### Sampling Observations
Notes: we might want to look at a small sample of households for more detail

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(450)
sample.ids<-sample(levels(yo$id),16)

ggplot(data=subset(yo,id %in%  sample.ids),
       aes(x=time,y=price))+
  facet_wrap(~id)+
  geom_line()+
  geom_point(aes(size=all.purchases),pch=1)

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r}
library(GGally)
set.seed(1836)
pf_subset<-pf[,c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset),1000),])
```

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

# QUIZ proportion of friendships initiated

Your task is to create a new variable called 'prop_initiated' in the Pseudo-Facebook data set. The variable should contain
the proportion of friendships that the user initiated.


```{r create proportion friendships initiated}
suppressWarnings(suppressMessages(library(dplyr)))
pf %>%
  mutate(prop_initiated=friendships_initiated/friend_count)->pf
```


# Create a line graph of the median proportion of friendships initiated ('prop_initiated') vs. tenure and color the line segment by year_joined.bucket.



```{r median proportion initiated by tenure}
ggplot(data=subset(pf,friend_count>0),
       aes(x=tenure,y=prop_initiated,color=year_joined.bucket))+
  geom_line(stat='summary',fun.y=median)
```

## Smoothed version

```{r median proportion initiated by tenure smoothed}
ggplot(data=subset(pf,friend_count>0),
       aes(x=30 * round(tenure/30),
           y=prop_initiated,color=year_joined.bucket))+
  geom_line(stat='summary',fun.y=median)
```

Mean proportion of friendships initiated by year_joined.bucket


```{r mean proportion initiated by year joined bucket}
pf%>%
  filter(friend_count>0)%>%
  group_by(year_joined.bucket)%>%
  summarise(mean(prop_initiated))
```


In theory, if we assume that people initate the most friendships when the first join, newer users have much more users that they can initiate friendships with.







